\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-27 paper.qmd
Updated Figure 2 and 3 with their titles

sc_ltr_pos <- c(0.08, 0.9)
# sc_xlims <- c(-0.5, 1.47)
# sc_ylims <- c(-0.32, 2.1)
sc_xlims <- c(-0.3, 1.3)
sc_ylims <- c(-0.3, 3.2)

nldr_two_curvy <- umap_two_curvy_scaled |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos)
```

@@

```{r}
#| label: hexbin-two-curvy
## Compute hexbin parameters
num_bins_x_two_curvy <- 7

@@

two_curvy_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#253494") +
  geom_point(
    #size = 0.2,
    alpha = 0.3,
    color = "#969696") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("a", cex = 2) +
  theme(
    legend.position = "none"
  )

proj_plot_two_curvy_true <- model_df |>
  ggplot(
    aes(
      x = proj1_from, 
      y = proj2_from)) +
  geom_point(
    color = "#000000") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("a", cex = 2) +
  theme(
    legend.position = "none"
  )

@@

two_curvy_proj_tsne_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#253494") +
  geom_point(
    #size = 0.2,
    alpha = 0.3,
    color = "#969696") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.125, 0.075)) + 
  ylim(c(-0.125, 0.075)) +
  interior_annotation("b", c(0.08, 0.93), cex = 2) +
  theme(
    legend.position = "none"
  )
```

```{r}
#| label: fig-two-curvy-true-proj
#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two-curvy data (black points). The data is spread along the S shape, and does not vary much from this curve. (The **langevitour** software is used to view the data with a tour, and the full video is available at <https://youtu.be/I5GL23vLiw0>)." 
#| fig-pos: H
#| fig-width: 10
#| fig-height: 5

two_curvy_proj_tsne_model1 + two_curvy_proj_tsne_model2 + 
  plot_layout(ncol = 2)
```

```{r}
#| label: fig-NLDR-two-curvy
#| echo: false
#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The S-curve data is shown."
#| fig-width: 8
#| fig-height: 4
#| out-width: 100%

@@

plot_predict_umap_two_curvy <- ggplot(
  data = jittered_points_df, 
  aes(x = UMAP1,
      y = UMAP2)) +
  geom_point(alpha = 0.5) +
  interior_annotation("c", c(0.92, 0.96), cex = 2)
```

@@

nldr_two_curvy + 
  trimesh_two_curvy_removed1_with_data +
  two_curvy_proj_umap_model +
  two_curvy_proj_umap_all_model +
  plot_layout(ncol = 2)
```

@@

plot_predict_umap_fun_two_curvy <- ggplot(
  data = predict_umap_two_curvy_scaled, 
  aes(x = UMAP1, 
      y = UMAP2)) +
  geom_point(alpha = 0.5) +
  interior_annotation("b", c(0.92, 0.96), cex = 2)
```

@@

plot_predict_umap_connect_two_curvy <- ggplot(
  data = predict_point_df, 
  aes(x = UMAP1, 
      y = UMAP2, 
      group = ID)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.3) +
  interior_annotation("d", c(0.92, 0.96), cex = 2)
```

>>>>

sc_ltr_pos <- c(0.08, 0.9)
# sc_xlims <- c(-0.5, 1.47)
# sc_ylims <- c(-0.32, 2.1)
sc_xlims <- c(-0.27, 1.2)
sc_ylims <- c(-0.2, 0.8)

nldr_two_curvy <- umap_two_curvy_scaled |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos)

nldr_two_curvy_original <- umap_two_curvy_scaled |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
  interior_annotation("a", sc_ltr_pos)
```

@@

```{r}
#| label: hexbin-two-curvy
## Compute hexbin parameters
num_bins_x_two_curvy <- 9

@@

two_curvy_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#000000") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("a", c(0.08, 0.93), cex = 2) +
  theme(
    legend.position = "none"
  )

proj_plot_two_curvy_true <- model_df |>
  ggplot(
    aes(
      x = proj1_from, 
      y = proj2_from)) +
  geom_point(
    color = "#000000") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("a", c(0.08, 0.93), cex = 2) +
  theme(
    legend.position = "none"
  )

@@

two_curvy_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#000000") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.125, 0.075)) + 
  ylim(c(-0.125, 0.075)) +
  interior_annotation("b", c(0.08, 0.93), cex = 2) +
  theme(
    legend.position = "none"
  )
```

```{r}
#| label: fig-two-curvy-true-proj
#| fig-cap: "Two views of the true model (blue lines) in $2\\text{-}D$ projections from $7\\text{-}D$, for the two-curvy data (black points). The data is spread along the non-linear shaped clusters, and does not vary much from the curves. The **langevitour** software is used to view the data with a tour, and the full video is available at <>." 
#| fig-pos: H
#| fig-width: 10
#| fig-height: 5

two_curvy_proj_umap_model1 + two_curvy_proj_umap_model2 + 
  plot_layout(ncol = 2)
```

```{r}
#| label: fig-NLDR-two-curvy
#| echo: false
#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$): (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids. The S-curve data is shown."
#| fig-width: 12
#| fig-height: 4
#| out-width: 100%

@@

plot_predict_umap_two_curvy <- ggplot(
  data = jittered_points_df, 
  aes(x = UMAP1,
      y = UMAP2)) +
  geom_point(alpha = 0.5) +
  interior_annotation("c", c(0.08, 0.93), cex = 2)
```

@@

nldr_two_curvy_original + 
  trimesh_two_curvy_removed1_with_data +
  two_curvy_proj_umap_model +
  two_curvy_proj_umap_all_model +
  plot_layout(ncol = 2)
```

@@

plot_predict_umap_fun_two_curvy <- ggplot(
  data = predict_umap_two_curvy_scaled, 
  aes(x = UMAP1, 
      y = UMAP2)) +
  geom_point(alpha = 0.5) +
  interior_annotation("b", c(0.08, 0.93), cex = 2)
```

@@

plot_predict_umap_connect_two_curvy <- ggplot(
  data = predict_point_df, 
  aes(x = UMAP1, 
      y = UMAP2, 
      group = ID)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.3) +
  interior_annotation("d", c(0.08, 0.93), cex = 2)
```

### 2024-09-27 paper.tex
Updated Figure 2 and 3 with their titles

\caption{\label{fig-two-curvy-true-proj}Two views of the true model
(blue lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
two-curvy data (black points). The data is spread along the S shape, and
does not vary much from this curve. (The \textbf{langevitour} software
is used to view the data with a tour, and the full video is available at
\url{https://youtu.be/I5GL23vLiw0}).}

>>>>

\caption{\label{fig-two-curvy-true-proj}Two views of the true model
(blue lines) in \(2\text{-}D\) projections from \(7\text{-}D\), for the
two-curvy data (black points). The data is spread along the non-linear
shaped clusters, and does not vary much from the curves. The
\textbf{langevitour} software is used to view the data with a tour, and
the full video is available at \textless\textgreater.}
